---
title: "Lab 5"
author: "Rahmat Ashari"
date: " `r format(Sys.time(), format = '%b %d, %Y')` "
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1
```{r}

getwd()

```

# Task 2
```{r}

mybin=function(iter,n,p, tabel_only = F){ 
  
  # make a matrix to hold the samples
  #initially filled with NA's
  sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
  #Make a vector to hold the number of successes in each trial
  succ=c()
  
  for( i in 1:iter){
    #Fill each column with a new sample
    sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))
    #Calculate a statistic from the sample (this case it is the sum)
    succ[i]=sum(sam.mat[,i])
  }
  
  #Make a table of successes
  succ.tab=table(factor(succ,levels=0:n))
  #Make a barplot of the proportions
  barplot(succ.tab/(iter), col=rainbow(n+1), xlab="Number of successes")
  succ.tab/iter
  
  if (tabel_only == T){
    dev.off()
    succ.tab
  } 
  
}

#mybin(iter = 100, n = 10, p =0.7)
trial = c(100, 200, 500, 1000, 10000)
for (i in trial) {
  mybin(i, n = 10, p = 0.7)
  tex = paste('Binomial Distribution Simulation with p = 0.7 & Iterations =',as.character(i) )
  title(tex)
}

tableSucc = mybin(10000, 10, 0.7, tabel_only = T)
tableSucc
yeet = as.data.frame(tableSucc)
toShow = yeet[,2]

#help("dbinom")
tr = dbinom(1:11, 11, prob = 0.7)*10000

#check simulation performace
check = matrix(1:22, nr = 2, byrow = T)
check[1,] = tr
check[2,] = toShow
row.names(check) = c('dbinom', 'simulation')
check

```

Results from simulation and  dbinom are arguably close.

# Task 3
```{r}

#20 marbles, 12 whites (denoted by 1) and 8 blacks (denoted by 0)
marbles = c(1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,0)

#sampling without replacement
sample(marbles, size = 5, replace = F)

#sampling with replacement
sample(marbles, size = 5, replace = T)

myhyper = function(iter=100,N=20,r=12,n=5, label_only = F){
  
  # make a matrix to hold the samples
  #initially filled with NA's
  sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
  #Make a vector to hold the number of successes over the trials
  succ=c()
  
  for( i in 1:iter){
    #Fill each column with a new sample
    sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
    #Calculate a statistic from the sample (this case it is the sum)
    succ[i]=sum(sam.mat[,i])
  }
  
  #Make a table of successes
  succ.tab=table(factor(succ,levels=0:n))
  #Make a barplot of the proportions
  barplot(succ.tab/(iter), col=rainbow(n+1), xlab="Number of successes")
  succ.tab/iter
  
  if (label_only == T){
    dev.off()
    succ.tab
  }
}

for (i in trial) {
  myhyper(iter = i, N=20,r=12,n=5, label_only = F)
  tex = paste('HYPERGEOMETRIC Simulation with Iterations =',as.character(i) )
  title(tex)
}

tableSuccH = myhyper(10000, N=20,r=12,n=5, label_only = T)
tableSuccH

hyperval = as.data.frame(tableSuccH)
hyperval_show = hyperval[,2]

trr = dhyper(1:6, 12, 8, 6)*10000

#check simulation performace
check2 = matrix(1:12, nr = 2, byrow = T)
check2[1,] = trr
check2[2,] = hyperval_show
row.names(check2) = c('dhyper', 'simulation')
check2

```

Values from simulation and dhyper are close, but not as great as binomial simuation. 

# Task 4
```{r}

mysample = function(n, iter=10,time=0.5, one_only = F){ 
  
  #the addition of one_only input is to prevent the later sub task to show 30 plots. 
  
  if (one_only == F){
    for( i in 1:iter){
      #make a sample
      s=sample(1:10,n,replace=TRUE)
      # turn the sample into a factor
      sf=factor(s,levels=1:10)
      #make a barplot
      barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
              main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
              ylim=c(0,0.2)
      )
      
      #release the table
      Sys.sleep(time)
    }
  } else {
    s=sample(1:10,n,replace=TRUE)
    sf=factor(s,levels=1:10)
    barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
              main=paste("Example sample()", " random iteration ", " n= ", n,sep="") ,
              ylim=c(0,0.2))
  }
  
}

```

The function mysample() generates n number of samples between the numeric 1 and 10, and plot their frequency iteration number of times. 

```{r}

mysample(n=1000, iter=30,time=1, one_only = T)

```

Each bin seems to have equal weight for each iteration. 

# Task 5
Calculate $$\binom{8}{4}$$
```{r}

choose(8,4)

```


Compute $$P(Y > 4), Y \sim Pois(\lambda = 2)$$
```{r}

res_pois = c()

for (i in 0:4){
  p = 2^i * (exp(-2)) / factorial(i)
  res_pois = append(res_pois, p)
}

1 - sum(res_pois)
```


Compute $$P(Y = 10), Y \sim NegBin(p=0.4, r = 3)$$
```{r}

dnbinom(7,3,0.4)

```



Compute $$P(Y \leq 8), Y \sim Bin(n = 15, p =0.4)$$
```{r}

res = c()

for (i in 0:8){
  bino = choose(15, i)
  cp = bino * (0.4^i) * ((1-0.4)^(15-i))
  res = append(res, cp)
}


sum(res)
```

# Task 6
```{r}

library(math4753)

#test using function with  default input values
mybin(iter=100,n=10, p=0.5)

```


