---
title: "Lab 7"
author: "Rahmat Ashari"
date: " `r format(Sys.time(), format = '%b %d, %Y')` "
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Task 1
```{r}

getwd()

```

# Task 2

### Generate chi sq statistics with different sample size 
```{r}


mychisim <- function(n1=10, sigma1=3, mean1=5, iter=1000, ymax=0.1, xleg = 0, ...){ # adjust ymax to make graph fit
  
  y1 = rnorm(n1*iter, mean=mean1, sd=sigma1) # generate iter samples of size n1 for each iteration
  # y1 is a vector of numerics
  
  data1.mat = matrix(y1, nrow=n1, ncol=iter, byrow=TRUE) # create a matrix with column representing an iter
  
  ssq1 = apply(data1.mat,2,var) # get variance of each iter
  #ssq1 is a vector
  
  w = (n1-1)*ssq1/sigma1^2 # chi-sq stat
  # w will also be a vector
  
  hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
       main=substitute(paste("Sample size = ",n[1]," = ",n1," statistic = ",chi^2)),
       xlab=expression(paste(chi^2, "Statistic",sep=" ")), las=1)
  
  lines(density(w),col="Blue",lwd=3) # add a density plot of w
  
  curve(dchisq(x,n1-1),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  
  title=expression(chi^2==frac((n[1]-1)*s^2,sigma^2)) #mathematical annotation -see ?plotmath
  
  legend(x = xleg, y = ymax - 0.01,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #

  invisible(list(w=w,summary=summary(w),sd=sd(w),fun="Chi-sq"))
  
  #print('Save the output of the function as an object to see statistical chi sq summary by indexing the 
        #object as a list.')
  
}

mychisim(n1 = 10, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.15, xleg = 20)
mychisim(n1 = 20, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.1, xleg = 30)
mychisim(n1 = 100, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.05, xleg = 120)
mychisim(n1 = 200, sigma1 = 4, mean1 = 10, iter = 1000, ymax = 0.04, xleg = 240)



```

### Histogram of chisq

```{r}

chisq = mychisim(n1 = 10, sigma1 = 10, mean1 = 20, iter = 1500, ymax = 0.15, xleg = 20)
hist(chisq$w, col = '#0DEADB')

```


# Task 3


### Generate T statistics with different sample size 
```{r}

myTsim <- function(n1=10, sigma1=3, mean1=5, iter=1000, xleg = 4, ymax=1,...){
  
  y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
  
  data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
  
  sd1=apply(data1.mat,2,sd) # sd
  
  ybar=apply(data1.mat,2,mean)  # mean
  
  w=(ybar-mean1)/(sd1/sqrt(n1))      #T stat
  
  hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
       main=substitute(paste("Sample size = ",n[1]," = ",n1," statistic = ",T," iterations= ",iter)),
       xlab=expression(paste(T, "Statistic",sep=" ")), las=1)
  
  lines(density(w),col="Blue",lwd=3) # add a density plot
  
  curve(dt(x,n1-1),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  
  title=expression(T==frac((bar(y)-mu),s/sqrt(n1))) #mathematical annotation -see ?plotmath
  
  legend(x = xleg, y = ymax - 0.1 ,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # 
  
  invisible(list(w=w,summary=summary(w),sd=sd(w),fun="T")) # some output to use if needed
  
}

myTsim(n1 = 10, iter = 1000, mean1 = 10, sigma1 = 4, xleg = 1, ymax=0.5)
myTsim(n1 = 20, iter = 1000, mean1 = 10, sigma1 = 4, xleg = 1, ymax=0.5)
myTsim(n1 = 100, iter = 1000, mean1 = 10, sigma1 = 4, xleg = 1, ymax=0.5)
myTsim(n1 = 200, iter = 1000, mean1 = 10, sigma1 = 4, xleg = 1, ymax=0.5)

```

### Histogram of T for a set of inputs
```{r}

T3 = myTsim(n1 = 10, iter = 1500, mean1 = 20, sigma1 = 10, xleg = 1.5, ymax=0.5)
hist(T3$w, col = '#0DEADB')

```

# Task 4

### Generate chi sq statistics from two populations
```{r}

mychisim2<-function(n1=10,n2=14,sigma1=3,sigma2=3,mean1=5,mean2=10,iter=1000, xleg = 12, ymax=0.07,...){    # adjust ymax to make graph fit
  
  y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
  
  y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
  
  data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
  
  data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
  
  ssq1=apply(data1.mat,2,var) # ssq1 is s squared
  
  ssq2=apply(data2.mat,2,var)
  
  spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
 
   w=(n1+n2-2)*spsq/(sigma1^2)#sigma1=sigma2,  Chi square stat
  
   hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
       main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",chi^2)),
       xlab=expression(paste(chi^2, "Statistic",sep=" ")), las=1)
  
   lines(density(w),col="Blue",lwd=3) # add a density plot
  
   curve(dchisq(x,n1+n2-2),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  
   title=expression(chi^2==frac((n[1]+n[2]-2)*S[p]^2,sigma^2)) #mathematical annotation -see ?plotmath
  
   #legend(locator(1),c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
   legend(x = xleg, y = ymax - 0.01 ,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)
   
   invisible(list(w=w,summary=summary(w),sd=sd(w),fun="Chi-sq")) # some output to use if needed
}

mychisim2(n1 = 10, n2 = 10, sigma1 = 4, sigma2 = 4, mean1 = 5, mean2 = 10, iter = 1000, xleg = 30)
mychisim2(n1 = 20, n2 = 10, sigma1 = 10, sigma2 = 10, mean1 = 3, mean2 = 5, iter = 1000, xleg = 40)
mychisim2(n1 = 50, n2 = 50, sigma1 = 4, sigma2 = 4, mean1 = 5, mean2 = 10, iter = 1000, xleg = 120, ymax = 0.04)
mychisim2(n1 = 80, n2 = 50, sigma1 = 10, sigma2 = 10, mean1 = 3, mean2 = 5, iter = 1000, xleg = 160, ymax = 0.04)

```

### Histogram of two-pop chi sq stats
```{r}

chisq2pop = mychisim2(iter = 10000, xleg = 40)
hist(chisq2pop$w, col = '#0DEADB')

```

# Task 5

### Create T stats function from 2 populations with legend location control
```{r}

myTsim2<-function(n1=10,n2=14,sigma1=3,sigma2=3,mean1=5,mean2=10,iter=1000,ymax=0.5,xleg =2,clickLeg = F,...){
  
  y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
  
  y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
  
  data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
  
  data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
  
  ssq1=apply(data1.mat,2,var) # ssq1 is s squared
  
  ybar1= apply(data1.mat,2,mean)
  
  ssq2=apply(data2.mat,2,var)
  
  ybar2=apply(data2.mat,2,mean)
  
  spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
  
  w=((ybar1-ybar2)-(mean1-mean2))/sqrt(spsq*(1/n1+1/n2))#sigma1=sigma2,  Chi square stat
  
  hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
       main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",T)),
       xlab=paste(" T Statistic",sep=""), las=1)
  
  lines(density(w),col="Blue",lwd=3) # add a density plot
  
  curve(dt(x,n1+n2-2),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  
  title=expression(T==frac((bar(Y)[1]-bar(Y)[2])-(mu[1]-mu[2]),S[p]*sqrt(frac(1,n[1])+frac(1,n[2])))) #mathematical annotation -see ?plotmath
  
  if (clickLeg == T){
    legend(locator(1),c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)# Legend #
  } else {
    legend(x = xleg, y = ymax - 0.1,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)# Legend #
  }
  invisible(list(w=w,summary=summary(w),sdw=sd(w),fun="T")) # some output to use if needed
}

```


### The Mathematcics of T statistics

The T statistics from normally distributed populations is expressed by

$$ T = \frac{(\overline{Y_1} - \overline{Y_2})-(\mu_1 - \mu_2)}
{S_p\sqrt{\frac{1}{n_1}+\frac{1}{n_2}}} $$

Where Y_i is the mean of the sampled observations, μ_i is the mean of the population, n_i is the size of the sample from each population, and Sp is the pooled sample variance.

From the code above, this expression is calculated by 

```
 w=((ybar1-ybar2)-(mean1-mean2))/sqrt(spsq*(1/n1+1/n2))
```

### Generate T stats from 2 populations
```{r}

myTsim2(n1 = 10, n2 = 10, mean1 = 5, mean2 = 10, sigma1 = 4, sigma2 = 4, iter = 1000, xleg = -4)
myTsim2(n1 = 20, n2 = 10, mean1 = 3, mean2 = 5, sigma1 = 10, sigma2 = 10, iter = 1000, xleg = -4)
myTsim2(n1 = 50, n2 = 50, mean1 = 5, mean2 = 10, sigma1 = 4, sigma2 = 4, iter = 10000, xleg = -4)
myTsim2(n1 = 80, n2 = 50, mean1 = 3, mean2 = 5, sigma1 = 10, sigma2 = 10, iter = 10000, xleg = -4)

```

### Histogram of two-pop T stats
```{r}

T5 = myTsim2(iter = 10000)
hist(T5$w, col = '#0DEADB')

```


# Task 6

### Defining F stats function from two populations
```{r}

myFsim2<-function(n1=10,n2=14,sigma1=3,sigma2=2,mean1=5,mean2=10,iter=1000,ymax=0.9,xleg = 6,...){

  y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
  y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
  data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
  data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
  ssq1=apply(data1.mat,2,var) # ssq1 is s squared
  ssq2=apply(data2.mat,2,var)
  #spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
  w=ssq1*sigma2^2/(ssq2*sigma1^2) #
  hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
       main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",F)),
       xlab=paste("F Statistic",sep=""), las=1)
  lines(density(w),col="Blue",lwd=3) # add a density plot
  curve(df(x,n1-1,n2-1),xlim=c(0,6),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  title=expression(F==frac(s[1]^2,s[2]^2)*frac(sigma[2]^2,sigma[1]^2)) #mathematical annotation -see ?plotmath
  legend(x = xleg, y = ymax - 0.1,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)# Legend #
  invisible(list(w=w,summary=summary(w),sd=sd(w),fun="F")) # some output to use if needed

}

```

### The Mathematics of F Statistics from two populations

$$ F = (\frac{S_1^2}{S_2^2})(\frac{\sigma_1^2}{\sigma_2^2}) $$

The assumption is that F is distributed with (n1 - 1) numerator degree of freedom and (n2 - 1) denominator degree of freedom. 


### Generate F stats from wwo populations
```{r}

myFsim2(n1 = 10, n2 = 10, sigma1 = 1, sigma2 = 2, mean1 = 10, mean2 = 13, iter = 1000, ymax = 1)
myFsim2(n1 = 15, n2 = 15, sigma1 = 1, sigma2 = 2, mean1 = 10, mean2 = 13, iter = 1000, ymax = 1, xleg = 3)
myFsim2(n1 = 15, n2 = 15, sigma1 = 2, sigma2 = 2, mean1 = 20, mean2 = 20, iter = 1000, ymax = 1, xleg = 3)
myFsim2(n1 = 15, n2 = 15, sigma1 = 1, sigma2 = 2, mean1 = 10, mean2 = 13, iter = 10000, ymax = 1, xleg = 3)
```


### Histogram of F stats
```{r}

F6 = myFsim2(iter = 10000)
hist(F6$w, col = '#0DEADB')

```

# Task 7

### Adding **data** into a package
```{r}

library(math4753)

data('fire')
knitr::kable(head(fire))

```

