---
title: "Assigment 3"
author: "Rahmat Ashari"
date: " `r format(Sys.time(), format = '%b %d, %Y')` "
output: 
  html_document:
    toc: yes
    toc_float: yes
---


# Question 1, MS 5.54
```{r}

data1 = read.csv("PHISHING.csv")
head(data1)
time1 = data1[,1]

```

## a
```{r}

#2 mins = 120 secs
1 - pgamma(120, shape = 1, scale = 95)

```

## b 
```{r}

# Sample mean 
mean(time1)

# Sample variance
var = sd(time1)^2
var

```

```{r}

# Exponential distribution
mean1 = 95
var1 = 95^2

## mean
mean1

## variance
var1

```

The means are extremely close, while the variances are arguably within close proximity. Overall, exponential distribution is sufficient in describing the phishing dataset. 

# Question 2, MS 5.56
```{r}

alpha = 3
beta = 0.07

```

## a
```{r}

# the mean
alpha * beta

# the variance
alpha * beta^2

```

## b 
```{r}

lower2 = 0.21 - (3*0.0147)
upper2 = 0.21 + (3*0.0147)

lower2
upper2

```

The maximum value observed 0.6 million cf/s is not within 3 std dev generated from the gamma distribution with $\alpha$ = 3 and $\beta$ = 0.07. Therefore, one or both of these parameters is/are wrong in describing the 4-year max flood level. 

# Question 3, MS 5.60

## a

Mean for formula A
```{r}

2*2

```

Mean for formula B
```{r}

1*4

```

## b 

variance for formula A
```{r}

2 * 2^2

```

variance for formula B
```{r}

1 * 4^2

```

## c
```{r}

# greater prob of generating reaction less than a min

forA = pgamma(1, shape = 2, scale = 2)
forB = pgamma(1, shape = 1, scale = 4)

forA
forB

```

Formula B has a greater probability of generating a human reaction in less than a minute. 

# Question 4, MS 5.74
```{r}
reparwbl = function(a,b){ # a,b from text
  list(ar=a,br=b^(1/a)) # ar, br for R
}
```

## a
```{r}

alpha4 = 2
beta4 = 4

reparwbl(alpha4,beta4)

ar = 2
br = 2

pweibull(2, 2, 2)

```

## b
```{r}

# mean
beta4^(1/alpha4) * gamma((alpha4 + 1)/ alpha4)

# standard dev
var4 = beta4^(2/alpha4)* (gamma((alpha4 + 2)/ alpha4) - (gamma((alpha4 + 1)/ alpha4)^2))

var4^0.5

```

## c
```{r}

lower4 = 1.7725 - (2*0.9265)
upper4 = 1.7725 + (2*0.9265)

lower4
upper4

pweibull(upper4, ar, br) - pweibull(lower4, ar, br)

```

## d
```{r}

1 - pweibull(6, ar, ar)

```

The probability in which Y>6 is really small: unlikely to exceed 6 years. 

# Question 5, MS 5.84
```{r}

# beta
alpha5 = 2
beta5 = 9
```

## a
```{r}
# mean
round(alpha5/(alpha5 + beta5), 4)

# variance
round((alpha5*beta5)/((alpha5 + beta5)^2 * (alpha5 + beta5 + 1)), 4)

```

## b 
```{r}

round(1 - pbeta(0.4, alpha5, beta5),4)

```

## c
```{r}

round(pbeta(0.1, alpha5, beta5), 4)

```


# Question 6, MS 5.114

## a

For Weibull distribution in a non-zero domain, 
$$ f(y)= \frac{\alpha}{\beta}y^{\alpha - 1}e^{-y^{\alpha}/\beta} $$

Matching the above equation with 
$$f(y) = \frac{1}{8}ye^{-y^{2}/16} $$

We can deduce that $\alpha$ = 2 and $\beta$ = 16

## b 
```{r}

alpha6 = 2
beta6 = 16

# mean
round(beta6^(1/alpha6) * gamma((alpha6 + 1)/alpha6),4)

# variance
round(beta6^(2/alpha6)* (gamma((alpha6 + 2)/ alpha6) - (gamma((alpha6 + 1)/ alpha6)^2)), 4)

```

## c
```{r}

reparwbl(alpha6, beta6)

ar6 = 2
br6 = 4

round(1 - pweibull(6, ar6, br6),4)

```

# Question 7, MS 6.2

## a

$$ \sum_{y}\sum_{x}p(x,y)=1$$

```{r}
library(MASS)

# two dice have 36 outcomes with equal likelyhood.

# joint probability distribution p(x,y) can be represented as table

joint = matrix(1:36, nrow = 6, byrow = T)
joint[,] = 1/36

colnames(joint) = c(1,2,3,4,5,6)
rownames(joint) = c(1,2,3,4,5,6)

fractions(joint)

```

## b 

$$P_{1}(x)=\sum_{y}p(x,y) $$
```{r}

#for p1(x)

fractions(sum(joint[1,]))

# for p2(y)

fractions(sum(joint[,1]))

```

## c
$$p_{1}(x|y) = \frac{p(x,y)}{p_2(y)} $$
```{r}

# from part a we know for any x and y, p(x,y) = 1/36

#for p1(x | y)
fractions((1/36)/(1/6))

# for p2(y | x)
fractions((1/36)/(1/6))

```

## d

From part **b** and **c**, since 
$$ p_{1}(x|y) = p_{1}(x)$$ 
(this is also true for $p_2$), the two dice events are independent. 

# Quetion 8, MS 6.4

## a

Given **X** representing the energy level and **Y** representing the time period,

```{r}

joint8 = matrix(1:9, nrow = 3, byrow = T)
colnames(joint8) = c(1,2,3)
rownames(joint8) = c(1,2,3)
# if we do by row, how many x_i correspond do y_i?
joint8[1,] = c(1/7, 2/7, 1/7)
joint8[2,] = c(0, 0, 2/7)
joint8[3,] = c(0, 0, 1/7)

# X is the column, Y is the row
fractions(joint8)

```

## b 
$$P_{1}(x)=\sum_{y}p(x,y) $$

```{r}

p1x8 = apply(joint8, 2, sum)
fractions(p1x8)

```

## c
$$P_{2}(y)=\sum_{x}p(x,y) $$

```{r}

p2y8 = apply(joint8, 1, sum)
fractions(p2y8)

```

## d
$$p_{2}(y|x) = \frac{p(x,y)}{p_1(x)} $$
```{r}

# for x = 1
joint8[,1]/p1x8[1]

# for x = 2
joint8[,2]/p1x8[2]

# for x = 3
joint8[,3]/p1x8[3]

```

# Question 9, MS 6.12

## a

$$ f(x,y) = \frac{e^{-y/10}}{10y}, 0<y<x<2y $$
$$f(y)=\int_{-\infty}^{\infty}\frac{e^{-y/10}}{10y}dx$$
In non-zero domain, $$f(y)=\int_{y}^{2y}\frac{e^{-y/10}}{10y}dx$$
$$ f(y) =  \frac{e^{-y/10}}{10y}x\Big|_y^{2y}$$
$$ f(y) =  \frac{e^{-y/10}}{10}$$
This is an exponential distribution

## b

E(Y) = $\mu$ = 10 for exponential distribution.

# Question 10, MS 6.14

## a

$$ \int_{-\infty}^{\infty}\int_{-\infty}^{\infty}f(x,y)dydx=1 $$
$$ \int_{0}^{\infty}\int_{0}^{x}ce^{-x^2}dydx=1 $$
$$ c\int_{0}^{\infty}e^{-x^2}y\Big|_0^{x}dx=1 $$
$$ c\int_{0}^{\infty}xe^{-x^2}dx=1 $$
$$ c\lim_{d\rightarrow\infty} -\frac{e^{-x^2}}{2}\Big|_0^{d}=1 $$
$$c\lim_{d\rightarrow\infty}(\frac{-e^{-d^2}}{2} - \frac{-e^{-0^2}}{2})=1$$
$$c(0-(-\frac{1}{2}))=1$$
$$ c = 2$$

## b

Finding the marginal density for X
$$ f(x) = \int_{-\infty}^{\infty}2e^{-x^2}dy$$
$$ f(x) = \int_{0}^{x}2e^{-x^2}dy$$
$$ f(x) = 2e^{-x^2}y\Big|_0^{x}$$
$$ f(x) = 2xe^{-x^2} $$

Showing that the integral over the entire x is equal to 1. But, the domain in which $f(x)$ is non-zero is only between 0 and $\infty$

```{r}

fun10 <- function(x){
  2*x*exp(-(x^2))
}

integrate(fun10, 0, Inf)

```

## c
$$f_2(y|x)=\frac{f(x|y)}{f(x)}$$
$$f_2(y|x)=\frac{2e^{-x^2}}{2xe^{-x^2}}$$
$$f_2(y|x)=\frac{1}{x}$$

# Question 11, MS 6.51
```{r}

joint11 = matrix(1:9, nr = 3, byrow = T)
colnames(joint11) = c(-1, 0, 1)
rownames(joint11) = c(-1, 0, 1)
joint11[1,] = c(1/12, 2/12, 1/12)
joint11[2,] = c(2/12, 0, 2/12)
joint11[3,] = c(1/12, 2/12, 1/12)
fractions(joint11)

```



in relation to E(X), the p(x) are
```{r}

fractions(apply(joint11, 2, sum))

```
Equal chance, E(X) = 0

in relation to E(Y), the p(y) are
```{r}

fractions(apply(joint11, 1, sum))

```
Equal chance, E(Y) = 0

To find covariance, 
$$cov(x,y)=E(XY)-E(X)E(Y)$$
$$cov(x,y)=\sum_{x}\sum_{y}xyp(x,y)-0$$
To find that double sum,
```{r}

sum(-1*joint11[1,] + 0*joint11[2,] + 1*joint11[3,])

```

$$cov(x,y)=0$$

For independence, the following must be true for all x and y
$$ p(x,y) = p(x)p(y) $$
suppose we take x = -1 and y = 1, 
$$p(-1,1)=p(-1)p(1) $$
For the left hand side, we can find from the table
$$ \frac{1}{12} = \frac{1}{3}*\frac{1}{3} $$
$$\frac{1}{12} \neq \frac{1}{9} $$

Overall, covariance being zero does not translate to independence. 

# Question 12, MS 6.74

## a
Y ~ Unif(1,3)
We know by Central Limit Theorem (CLT) that given the size of the sample is big, $E(\overline{Y})$ = $\mu$ 
```{r}

#E(Y)
(1+3)/2

```

## b 
Again, CLT states that $Var(\overline{Y}) =\frac{\sigma^2}{n}$
```{r}

var12 = ((3-1)^2/12)/60
var12

sd12 = var12^0.5
```

## c
CLT suggests that the means are distributed normal as:
$$ \overline{Y} \sim N(\mu=2, \sigma^2=0.0056) $$

## d
```{r}

# meanY between 1.5 and 2.5
pnorm(2.5, 2, sd12) - pnorm(1.5, 2, sd12)

```

## e
```{r}

# meanY exceeds 2.2
round(1 - pnorm(2.2, 2, sd12),4)

```

# Question 13, MS 6.90

## a
By CLT, 
$$ \overline{x} \sim N(\mu=p, \sigma^2=\frac{p(1-p)}{n})$$
```{r}

p = 0.4*20 #chance of swordfish to have beyond-treshold mercury level

var13 = 8*(20-8)/20
var13

sd13 = var13^0.5
sd13

```
$$ \overline{x} \sim N(0.4, 0.012)$$

```{r}

# normal approx to binomial, prob of <2 has beyond-treshold mercury
pnorm(1.5, p, sd13)

```

## b 
```{r}

# more than 1/2 of 20 samples have high mercury level/
# half of 20 is 10, so we use 10.5 as normal bound

1 - pnorm(10.5, p, sd13)

```

## c
```{r}

# calculating part a using the actual binomial distribution
dbinom(1,20,0.4) + dbinom(0,20,0.4)

# calculating part b using the actual binomial distribution
list13 = c(0,1,2,3,4,5,6,7,8,9,10)

tempo13 = 0

for (i in list13){
  tempo13 = tempo13 + dbinom(i, 20, 0.4)
}

1- tempo13

```

# Question 14, MS 7.108

```{r}

data14 = read.csv("LEADCOPP.csv")
head(data14)
lead = data14[,1]
copper = data14[,2]

```

## a
```{r}

t.test(lead, conf.level = 0.99)

```

## b
```{r}

t.test(copper, conf.level = 0.99)

```

## c

**for part a**, the output of the confidence interval is not physical. We cannot have negative contamination level. Therefore, we can say that 99% of the mean lead contamination data are entrapped within 0 and 6.9191 $\mu g/L$. This is quite a wide range, but again, the confidence interval is also big. 

**for part b**, we can say that 99% of the mean copper contamination data are entrapped within 0.1519 and 0.6647 mg/L. This is quite a wide range, but again, the confidence interval is also big. 

## d

99% confidence interval can be interpreted quite literally that we are sure, by 99%, the mean of the data series of interest lies within this interval. Quantitatively, 99% of the data lies within this interval.

# Question 15, MS 7.114

```{r}

data15 = read.csv("SOLARAD.csv")
stjos = data15[,2]
iowa = data15[,3]

diff15 = stjos - iowa

t.test(diff15, conf.level = 0.95)

```

The mean difference lies with 95% confidence between 156.8193 and 239.1807 BTU/sq. ft. 

# Question 16, MS 7.116

```{r}

data16 = read.csv("DIAZINON.csv")

day = data16[,2]
night = data16[,3]

t.test(day, conf.level = 0.9)
t.test(night, conf.level = 0.9)

diff16 = day - night
t.test(diff16, conf.level = 0.9)

```

## b 

some assumptions for the validity: both populations are independent, are normal, and have equal variances 

## c

Yes, considering the negative upper and lower bounds of the interval from part **a**, the mean diazinon residue level differs from day to night. 
