---
title: "Lab 11"
author: "Rahmat Ashari"
date: " `r format(Sys.time(), format = '%b %d, %Y')` "
output: 
  html_document:
    toc: yes
    toc_float: yes
    in_header: 'header.html'
    css: 'index.css'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1
```{r}

getwd()

```

# Task 2, Calculating CI

A ram press makes washers for agricultural machinery. The mean diameter (cm) of the washers is of interest. We can manually construct various CI level for the mean using pivotal statistics *t*. 

```{r}

d=c(5.0581, 4.9707, 5.0893, 4.9334, 4.9777, 5.0285, 4.8555, 4.9565, 4.9769, 4.9722, 4.999, 4.9925, 4.9686, 5.0662, 4.9239, 4.9781, 5.0485, 5.0014, 4.9957, 5.0195, 5.0118, 4.9928, 5.0361, 5.0185, 4.9879)

alpha = c(0.05, 0.1, 0.2, 0.5)

for (i in alpha){
  tstat = qt(1 - (i/2), length(d) - 1)
  interval = tstat*sd(d)/(length(d))^0.5
  print(paste0('for confidence = ', 1-i, ' the CI is between ', round(mean(d)-interval,4), ' and ', round(mean(d)+interval,4)))
}

```

We can check the quality of this manual approach. Lets compute 80% CI using `t.test()` and see the bounds given in the output. 

```{r}

obj2 = t.test(d, conf.level = 0.8)

obj2$conf

```

The answers from both approaches are the same. 

We can also build CI for $\sigma^2$ using $\chi^2$

```{r}

n = length(d)

for (i in alpha){
  low = round((n-1)*var(d)/qchisq(1 - i/2, n-1), 4)
  up = round((n-1)*var(d)/qchisq(i/2, n-1), 4)
  print(paste0('variance with CI = ', 1-i, ' is between ', low, ' and ', up))
}

```

# Task 3, Difference Between Two Means

We have two vectors representing the lengths of two different fish. We will attempt to use **R** to compute different CIs for the mean difference. 

First, the 95% CI for $\mu_{snapper}-\mu_{blue}$ will be computed manually. We will use the pooled variance approach since the length of the vectors are < 30. 
```{r}

blue=c(21.65, 17.48, 20.1, 21.57, 14.82, 19.17, 21.08, 18.23, 22.93, 15.66, 20.89, 21.66, 18.5, 20.59, 18.63, 18.91, 19.53, 17.7, 16.5, 19.03)

nb = length(blue)

snapper=c(31.65, 27.48, 30.1, 31.57, 24.82, 29.17, 31.08, 28.23, 32.93, 25.66, 30.89, 31.66, 28.5, 30.59, 28.63)

ns = length(snapper)

# pooled variance
pvar = (((ns - 1)*var(snapper)) + ((nb - 1)*var(blue)))/(ns+nb-2)

# t statistics
tstat3 = qt(1 - (0.05/2), ns+nb-2)

# distance from the mean 
dis = tstat3*(pvar*((1/ns)+(1/nb)))*0.5

# sample diff
mudiff = mean(snapper) - mean(blue)

# interval
print(paste0(mudiff - dis, ' ', mudiff + dis))

```

From repetitions of sampling, 95% of the mean difference is negative, suggesting that the means are indeed, different. 

We can also use `t.test()` to calculate the mean difference, assuming the variances of the two samples are the same. 

```{r}

ci3 = c(0.95, 0.85, 0.75, 0.25)

for (i in ci3){
  obj3 = t.test(snapper,blue,var.equal=TRUE, conf.level = i)
  print(obj3$conf.int)
}

```

As the confident level decreases, the confident interval shrinks, enclosing the point estimate tighter. 

# Task 4, Two Paired, Dependent Samples

Here, we calculate the mean difference for a paired samples, Exam 1 and Exam 2. First, we will compute the mean difference manually for a confidence level of 95%. 

```{r}

exam1=c(40.98, 59.36, 46.69, 41.8, 61.63, 65.31, 62.96, 60.21, 56.89, 78.41, 53.44, 75.2, 60.54, 52.43, 41.41, 70.79, 73.55, 55.65, 61.43, 63.84, 58.07, 53.79, 54.45, 67.18, 44.46)

exam2=c(50.22, 66.19, 58.75, 51.88, 66.61, 70.86, 74.25, 70.23, 69.55, 87.18, 63.62, 81.7, 70.5, 66.02, 51.35, 80.92, 85.65, 65.44, 74.37, 75.28, 67.86, 59.92, 64.42, 73.57, 57.15)

diff = exam1 - exam2
meandiff = mean(diff)
sddiff = sd(diff)
n4 = length(exam1) # this will be the same as length(exam 2) bcs of paired samples
zstat = qnorm(1 - (0.05/2), 0, 1)

interval4 = zstat*sddiff/n4^0.5

low4 = round(meandiff - interval4, 4)
up4 = round(meandiff + interval4, 4)

print(paste0('Mean diff for the paired sample with 95% conf level is ', low4, ' ', up4))

```

This means, 95% of repeating samples will lie within such interval, which is both lower and upper bounds < 0. Hence, we can say we are 95% confident that the means are different. 


Now we can use `t.test()` for various confidence levels.
```{r}

ci4 = c(0.9, 0.8, 0.7, 0.6, 0.1)

for (i in 1:5){
  obj4 = t.test(exam1,exam2, paired =T, conf.level = ci4[i])
  print(obj4$conf.int)
}

```

# Task 5, CI for a Population Proportion

We are given a data:

```{r}

data5 = read.csv('NZBIRDS.csv')
with(data5, table(Extinct,Flight))

```

find a 95% CI for the difference in proportion of flightless birds for extinct and non-extinct species.

```{r}

# Extinct
pe = 21/38
qe = 1-pe
ne = 38

# Not extinct
pn = 7/78
qn = 1 - pn
nn = 78

pe*ne
qe*ne
pn*nn
qn*nn
```

All of the results above, $p_in_i$ and $q_in_i$ are greater than 4. Hence, normal approximation is ok!

```{r}
# mean prop
meandiff5 = pe - pn

# z stat
zstat5 = qnorm(1 - (0.05/2), 0, 1)

dis5 = zstat5*((pe*qe/ne)+(pn*qn/nn))^0.5

low5 = meandiff5 - dis5
up5 = meandiff5 + dis5

low5
up5


```

# Task 6, Variance Ratio

First, we will evaluate the CI for the variance ratio between two random samples as seeded below using F statistics and 95% confidence level. 

```{r}

set.seed(35)
sam1=rnorm(25,mean=10,sd=5)
set.seed(45)
sam2=rnorm(34,mean=40,sd=8)

ratio = var(sam1)/var(sam2)

df1 = length(sam1) - 1
df2 = length(sam2) - 1

fstat = qf(1 - (0.05/2), df1, df2)

low6 = ratio*(1/fstat)
up6 = ratio*fstat

ratio
low6
up6

```

Now, we can evaluate the ratio of variance using `var.test()` with various confidence levels

```{r}

ci6 = c(0.8, 0.7, 0.6, 0.5)

for (i in ci6){
  obj6 = var.test(sam1, sam2, conf.level = i)
  print(obj6$conf.int)
}

```

# Task 7, Building a 95%-CI Function 

This section will build a function that calculates a 95% CI from a single vector x as the input. We will also add this function to my package. 

```

myci = function(x){
  n = length(x)
  tstat = qt(1 - (0.05/2), n-1)
  interval = tstat*sd(x)/(n)^0.5
  low = mean(x) - interval
  up = mean(x) + interval
  output = list('point est' = mean(x), 'low int' = low, 'up int' = up)
  return(output)
}

```

Now we can test the function from the package!

```{r}

library(math4753)

set.seed(23)
x = rnorm(30,mean=10,sd=12)

myci(x)

```

